15th International Satisfiability Modulo Theories Competition (SMT-COMP 2021): Rules and Procedures

### 2021-05-06 rules/rules21.tex
udpating text about old overall ranking change
updating rules with new divisions change

SMT-COMP 2021 will have four tracks: the \maintrack (before 2019: Main Track),
the \inctrack (before 2019: Application Track), the \ucoretrack, and the
\mvaltrack.
%
Within each track there are multiple divisions, where each division
uses benchmarks from a specific SMT-LIB logic (or group of logics).
We will recognize winners as measured by number of benchmarks solved
(taking into account the weighting detailed in
Section~\ref{sec:scoring}); we will also recognize solvers based on
additional criteria.

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  Sylvain Conchon, David D{\'e}harbe, Morgan Deters, Alberto Griggio,
  Liana Hadarean,
  Matthias Heizmann, Aina Niemetz, Albert Oliveras, Giles Reger, Aaron Stump,
  and Tjark Weber.}  describes the rules and competition procedures for
SMT-COMP~2021.
%
As in previous years, we have revised the rules slightly.  Some rule
changes in 2021 are designed to not overly punish solvers producing
syntactically wrong output, by treating those outputs as unknown and
not as unsoundness.  Moreover, we added more divisions to the model
validation track.
%
The principal changes from the previous competition rules are the following:
\begin{itemize}
  \item {\bf The option \akey{:print-success} is explicitly set to
    true or false.}  Previous competitions set this option to
    \akey{true} for the incremental track.  In addition, this
    competition will set the option to \akey{false} for all other
    tracks.
    \rationale{According to the standard the default value of
      \akey{:print-success} is true.  So a conforming solver would
      unnecessarily print \akey{success} after each command that needs
      to be stored and skipped by the post-processors.  Additionally,
      solvers can use this option as indication whether incremental
      checking is desired.}

@@

The \maintrack track will consist of selected non-incremental benchmarks in
each of the competitive logic divisions.  Each benchmark will be presented to
the solver as its first command-line argument.  The solver is then expected to
report on its standard output channel whether the formula is satisfiable
(\texttt{sat}) or unsatisfiable (\texttt{unsat}).  A solver may also report
\texttt{unknown} to indicate that it cannot determine satisfiability of the
formula.

@@

\header{Divisions.}
Within each track there are multiple divisions, and each division selects
benchmarks from a specific SMT-LIB logic in the SMT-LIB benchmark library.

@@

\header{Benchmark demographics.}
The set of all SMT-LIB benchmarks in a given division can be naturally
partitioned to sets containing benchmarks that are similar from the user
community perspective.  Such benchmarks could all come from the same
application domain, be generated by the same tool, or have some other
obvious common identity.
%
The organizers try to identify a meaningful partitioning based on the
directory hierarchy in SMT-LIB.  In many cases the hierarchy consists of
the top-level directories each corresponding to a submitter, who has
further imposed a hierarchy on the benchmarks.
%
The organizers believe that the submitters have the best information on
the common identity of their benchmarks and therefore partition each
division based on the bottom-level directory imposed by each submitter.
These partitions are referred to as \emph{families}.

\header{Benchmark selection.} \label{benchmark-selection}
The competition will use a large subset of SMT-LIB benchmarks, with some
guarantees on including new benchmarks.  In \textbf{all} tracks
% \textbf{except} the \challtrack,
the following selection process will be used.
\begin{enumerate}
\item \emph{Remove inappropriate benchmarks.} The
  organizers may remove benchmarks that are deemed inappropriate or
  uninteresting for competition, or cut the size of certain benchmark
  families to avoid their over-representation.  SMT-COMP attempts to
  give preference to benchmarks that are ``real-world,'' in the sense
  of coming from or having some intended application outside SMT.
\item \emph{Remove easy / uninteresting benchmarks.}
  For the following tracks, all benchmarks that can be
  considered as easy or uninteresting based on the following criteria
  will be removed.
  \begin{itemize}
    \item \emph{\maintrack.} All benchmarks that were solved by all
      solvers (including non-competitive solvers) in less than one second in
          the corresponding track in 2018, 2019, and 2020.
    \item \emph{\ucoretrack.} All benchmarks with only a single assertion.
  \end{itemize}
\item \emph{Cap the number of instances in a division.}
  The number of benchmarks in a division based on the size of the
  corresponding logic in SMT-LIB will be limited as follows.
  Let $n$ be the number of benchmarks in an SMT-LIB logic, then
  \begin{enumerate}
  \vspace{-1ex}
    \item \label{bench-sel-300} if $n \le 300$, all instances will be selected;
  \item \label{bench-sel-600} if $300 < n \leq 750$, a subset of 300 instances
    from the logic will be selected;
  \item \label{bench-sel-more} and if $n > 750$,
      40\% of the benchmarks of the logic will be selected.
  \end{enumerate}
\end{enumerate}
%
The selection process in cases \ref{bench-sel-600} and \ref{bench-sel-more}
above will guarantee the inclusion of new benchmarks by first picking randomly
one benchmark from each new benchmark family.  The rest of the benchmarks will
be chosen randomly from the remaining benchmarks using a uniform distribution.
%
The benchmark selection script will be publicly available at
\url{https://github.com/SMT-COMP/smt-comp/tree/master/tools} and
will use the same random seed as the rest of the competition.  The set of
benchmarks selected for the competition will be published when the competition
begins.

@@

For each track and division, we compute a division score based on the parallel
performance of a solver (the \emph{parallel division score}).  For the
\maintrack, \ucoretrack and \mvaltrack we also compute a division
score based on the sequential performance of a solver (the \emph{sequential
division score}).  Additionally, for the \maintrack, we further determine three
additional scores based on parallel performance: The \emph{24-second score}
will reward solving performance within a time limit of 24 seconds (wall clock
time), the \emph{sat score} will reward (parallel) performance on satisfiable
instances, and the \emph{unsat score} will reward (parallel) performance on
unsatisfiable instances.

@@

This year we will replace the competition-wide score with two new \emph{rankings} that select one solver per division and then rank those solvers. The rationale here is to take the focus away from the number of divisions entered and focus on measures that make sense to use to compare different divisions.

>>>>

SMT-COMP 2021 will have four tracks: the \maintrack (before 2019: Main Track),
the \inctrack (before 2019: Application Track), the \ucoretrack, and the
\mvaltrack.
%
Within each track there are multiple divisions, where each division
uses benchmarks from a specific group of SMT-LIB logics.
We will recognize winners as measured by number of benchmarks solved
(taking into account the weighting detailed in
Section~\ref{sec:scoring}); we will also recognize solvers based on
additional criteria.

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  Sylvain Conchon, David D{\'e}harbe, Morgan Deters, Alberto Griggio,
  Liana Hadarean,
  Matthias Heizmann, Aina Niemetz, Albert Oliveras, Giles Reger, Aaron Stump,
  and Tjark Weber.}  describes the rules and competition procedures for
SMT-COMP~2021.
%
As in previous years, we have revised the rules slightly.  Some rule
changes in 2021 are designed to not overly punish solvers producing
syntactically wrong output, by treating those outputs as unknown and
not as unsoundness.  Moreover, we added more divisions to the model
validation track.
%
The principal changes from the previous competition rules are the following:
\begin{itemize}
  \item {\bf \maintrack.} Previously, every logic in the \maintrack was its own
  division.  There are currently over 60 logics in SMT-LIB, which makes a short
  summary of the results impossible.  On the other hand the competition-wide
  scoring does a comparison between solvers that cannot be sensibly compared as
  they support completely disjoint logics. To get a better overview, this year
  we have in the \maintrack divisions combining related logics, thus reducing
  the number of divisions and obtaining more manageable and meaningful results.
  Rather than logics, solvers will declare the divisions they enter in the
  \maintrack. They will be ranked by the number of benchmarks they solve across
  all logics in the divisions entered.
  \begin{itemize}
    \item Quantifier-free divisions:
    \begin{itemize}
      \item QF\_Equality\\
Logics: QF\_UF, QF\_AX, QF\_DT
\item QF\_Equality+LinearArith\\
Logics: QF\_ALIA, QF\_AUFLIA, QF\_UFLIA, QF\_UFLRA, QF\_UFIDL
\item QF\_Equality+NonLinearArith\\
Logics: QF\_UFNRA, QF\_UFNIA, QF\_ANIA, QF\_AUFNIA
\item QF\_Equality+BVArith\\
Logics: QF\_ABV, QF\_UFBV, QF\_AUFBV
\item QF\_LinearIntArith\\
Logics: QF\_LIA, QF\_LIRA, QF\_IDL
\item QF\_LinearRealArith\\
Logics: QF\_LRA, QF\_RDL
\item QF\_BV
\item QF\_FPArith\\
Logics: QF\_FP, QF\_UFFP, QF\_FPLRA, QF\_BVFP, QF\_ABVFP, QF\_BVFPLRA, QF\_ABVFPLRA
\item QF\_NonLinearIntArith\\
Logics: QF\_NIA, QF\_NIRA
\item QF\_NRA
\item QF\_Strings\\
Logics: QF\_S, QF\_SLIA

    \end{itemize}
    \item Divisions with quantifiers:
    \begin{itemize}
      \item Equality\\
Logics: UF, UFDT
\item Equality+LinearArith\\
Logics: ALIA, AUFLIA, UFLIA, UFIDL, AUFLIRA, UFLRA, UFDTLIA, UFDTLIRA, AUFDTLIA, AUFDTLIRA
\item Equality+MachineArith\\
Logics: AUFFPDTLIRA, UFFPDTLIRA, UFFPDTNIRA, ABVFP, ABVFPLRA, UFBV, AUFBVDTLIA
\item Equality+NonLinearArith\\
Logics: AUFDTNIRA, UFDTNIA, UFDTNIRA, AUFNIA, AUFNIRA, UFNIA
\item Arith\\
Logics: LRA, LIA, NIA, NRA
\item BV
\item FPArith\\
Logics: BVFP, FP, BVFPLRA, FPLRA
    \end{itemize}
  \end{itemize}

  \rationale{We are aware that any partitioning of logics into divisions can be
    disadvantageous to some solvers that support some but not all logics in the
    division. We believe however that to a lesser extent this situation is already
    common, thinking of different features among problems in the same logic. So in
    our view the disadvantages are offset by the benefit of having a better
    presentation of the results. Regardless, results per logic will still be
    accessible within the results page of a given division, similarly to how one can
    now see the different scores other than sequential for a given logic.

    The divisions above were defined according to the following rationale:

    \begin{itemize}
      \item Similar logics that do not standout by themselves were combined.
      \item Integer and real, as well as bit-vector and floating-point,
      arithmetics are separated
      \begin{itemize}
        \item Note that the ``Arith'' division goes against this. However, its
        composing logics do not standout out by themselves as currently
        represented in SMT-LIB.
      \end{itemize}
      \item Arrays and datatypes are handled in a similar enough way to UF to be combined with it.
        \item Floating-point logics are combined since generally this is the dominating component for any logic contaning FP.
    \end{itemize}
  }

  \item {\bf The option \akey{:print-success} is explicitly set to
    true or false.}  Previous competitions set this option to
    \akey{true} for the incremental track.  In addition, this
    competition will set the option to \akey{false} for all other
    tracks.
    \rationale{According to the standard the default value of
      \akey{:print-success} is true.  So a conforming solver would
      unnecessarily print \akey{success} after each command that needs
      to be stored and skipped by the post-processors.  Additionally,
      solvers can use this option as indication whether incremental
      checking is desired.}

@@

The \maintrack track will consist of selected non-incremental benchmarks in
each of the competitive divisions.  Each benchmark will be presented to
the solver as its first command-line argument.  The solver is then expected to
report on its standard output channel whether the formula is satisfiable
(\texttt{sat}) or unsatisfiable (\texttt{unsat}).  A solver may also report
\texttt{unknown} to indicate that it cannot determine satisfiability of the
formula.

@@

\header{Divisions.}

Within each track there are multiple divisions, and each division selects
benchmarks from a specific group of SMT-LIB logics in the SMT-LIB benchmark
library.

@@

\header{Benchmark demographics.}
%
The set of all SMT-LIB benchmarks in the logics of a given division can be
naturally partitioned to sets containing benchmarks that are similar from the
user community perspective.
%
Such benchmarks could all come from the same
application domain, be generated by the same tool, or have some other
obvious common identity.
%
The organizers try to identify a meaningful partitioning based on the
directory hierarchy in SMT-LIB.  In many cases the hierarchy consists of
the top-level directories each corresponding to a submitter, who has
further imposed a hierarchy on the benchmarks.
%
The organizers believe that the submitters have the best information on the
common identity of their benchmarks and therefore partition each logic in a
division based on the bottom-level directory imposed by each submitter.  These
partitions are referred to as \emph{families}.

\header{Benchmark selection.} \label{benchmark-selection}
The competition will use a large subset of SMT-LIB benchmarks, with some
guarantees on including new benchmarks.  In \textbf{all} tracks
% \textbf{except} the \challtrack,
the following selection process will be used.
\begin{enumerate}
\item \emph{Remove inappropriate benchmarks.} The
  organizers may remove benchmarks that are deemed inappropriate or
  uninteresting for competition, or cut the size of certain benchmark
  families to avoid their over-representation.  SMT-COMP attempts to
  give preference to benchmarks that are ``real-world,'' in the sense
  of coming from or having some intended application outside SMT.
\item \emph{Remove easy / uninteresting benchmarks.}
  For the following tracks, all benchmarks that can be
  considered as easy or uninteresting based on the following criteria
  will be removed.
  \begin{itemize}
    \item \emph{\maintrack.} All benchmarks that were solved by all
      solvers (including non-competitive solvers) in less than one second in
          the corresponding track in 2018, 2019, and 2020.
    \item \emph{\ucoretrack.} All benchmarks with only a single assertion.
  \end{itemize}
\item \emph{Cap the number of instances in a division.}
  The number of benchmarks in a division based on the size of the
  corresponding logics in SMT-LIB will be limited as follows.
  Let $n$ be the number of benchmarks in an SMT-LIB logic, then
  \begin{enumerate}
  \vspace{-1ex}
    \item \label{bench-sel-300} if $n \le 300$, all instances will be selected;
  \item \label{bench-sel-600} if $300 < n \leq 750$, a subset of 300 instances
    from the logic will be selected;
  \item \label{bench-sel-more} and if $n > 750$,
      40\% of the benchmarks of the logic will be selected.
  \end{enumerate}
\end{enumerate}
%
The selection process in cases \ref{bench-sel-600} and \ref{bench-sel-more}
above will guarantee the inclusion of new benchmarks by first picking randomly
one benchmark from each new benchmark family.  The rest of the benchmarks will
be chosen randomly from the remaining benchmarks using a uniform distribution.
%
The benchmark selection script will be publicly available at
\url{https://github.com/SMT-COMP/smt-comp/tree/master/tools} and
will use the same random seed as the rest of the competition.  The set of
benchmarks selected for the competition will be published when the competition
begins.

@@

For each track and division, we compute a division score based on the parallel
performance of a solver (the \emph{parallel division score}).  For the
\maintrack, \ucoretrack and \mvaltrack we also compute a division
score based on the sequential performance of a solver (the \emph{sequential
division score}).  Additionally, for the \maintrack, we further determine three
additional scores based on parallel performance: The \emph{24-second score}
will reward solving performance within a time limit of 24 seconds (wall clock
time), the \emph{sat score} will reward (parallel) performance on satisfiable
instances, and the \emph{unsat score} will reward (parallel) performance on
unsatisfiable instances.
%
Finally, for the \maintrack all the above scores, in divisions composed by more
than one logic, will be presented not only for the overall division but also for
each logic composing the division.

@@

This score has been replaced the competition-wide score with two \emph{rankings} that select one solver per division and then rank those solvers. The rationale here is to take the focus away from the number of divisions entered and focus on measures that make sense to use to compare different divisions.
